\textls[-0]{\thetitle}

### 2023-03-25 canonicitymodel.tex
Done adding color for parsing.
cut clutter

\textbf{Unfortunately}, at this point, our notation is already highly densed because of frequent appearances of $\bullet$. These burden

Now we can extend the model above to include \TT constructs.
%We still start with the judgments.
%Since linkage transformers are just syntactic sugar, we will not interpret it.
The key idea of this extension is similar to that of the syntactic translation:
we interpret linkage types using the canonicity model of sigma types (developed in the prior work).
We use the inductive facility of the ambient meta\-logic to justify our W-types.

@@

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSigU{j}{\Gm}{\wsig}{n}} &: \GMWSig{j}{\denotesCC{\Gm}}{n} \\ 
    &\text{ i.e., a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  \text{we define }\TyLkg{}, \TyTkg{} & \text{ inductively on its input signature, via } \mathit{L},\Glued{\mathit{P}}  \\ 
  \mathit{L}("nil") &=  \denotesC{\top}  \\
  \mathit{L}((A,s,T) :: tl)(\gm)(\gm')(t) &= {\mathit{L}(tl)}(\gm)(\gm')(\lkgproj{1}{t}) \times \Gmodel{\GluedPi{A}{T}}(\gm)(\gm')(\lam{\lkgproj{2}{t}})   \\
  \Glued{\mathit{P}}("nil") &= ( \TyTkg{\LSigEmp} , \denotesC{\top} )\\
  \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{{\GSubstWeak{1}}}{s}}) \quad \text{(doing substitution on $T$)} \\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= {\mathit{L}(\denotesC{\lsig})} \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{P}}(\denotesC{\lsig})} \quad \text{ discard syntax info}  \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig} \quad \text{ given } \goodTerm{\Gm, \TyTkg{\lsig}}{s}{A} \\ 
  \denotesC{\goodSig{\Gm}{\sub{\lsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[0]} \quad \text{ take the first element in the tuple,...}\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= \Gmodel{("hd"~\denotesC{\lsig})[1]} \quad \text{  ... and discard syntax}\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[2]}\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
  \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesCC{\lkgproj{2}{\lkg}}}{\GSubstExt{{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
  \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_ \\ 
\end{align*}
This is the first point our proof becomes cryptic because of this bullet notation.

The first complexity comes from the three ``bullets after comma'' notation. The way to interpret this notation is that, this ``$\GSubstExt{\_}{\_}$'' should be considered as a binary function requiring two input of glued interpretation.

However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloading. The first one ``$\Gpair{\Glued{\ell'}}{...}$'' should be the glued interpretation for dependent tuple, but the other two should be the one for substitution extension.

Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution. Again, ``$\Gsub{\_}{\_}$'' should be considered as a binary function consuming two glued interpretation, corresponding to the substitution operation ``$\sub{\_}{\_}$''.

Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.

In all, the whole interpretation should be considered as:
\begin{enumerate}
  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
\end{enumerate}

>>>>

%\textbf{Unfortunately}, at this point, our notation is already highly densed because of frequent appearances of $\bullet$. These burden

Now we can extend the model above to include \TT constructs.
%We still start with the judgments.
%Since linkage transformers are just syntactic sugar, we will not interpret it.
The key idea of this extension is similar to that of the syntactic translation:
we interpret linkage types using the canonicity model of $\Sigma$-types (developed in the prior work).
We use the inductive facility of the ambient meta\-logic to justify our W-types.

@@

\begin{align*}
  \denotesC{\goodSig{\Gm}{\lsig}{n}} & \text{ is a list of 3-tuple of length $n$  } \\
  % & \text{ and we define  } {\goodSigC{\Gm}{\lsig}{n}} :\iff \denotesC{\goodSig{\Gm}{\lsig}{n}} \text{ is defined} \\ 
  \denotesC{\goodWSigU{j}{\Gm}{\wsig}{n}} &: \GMWSig{j}{\denotesCC{\Gm}}{n} \\ 
    &\text{ i.e., a list of 2-tuple of length $n$ } \\
  % & \text{ and we define } \goodWSigC{\Gm}{\wsig}{n} :\iff \denotesC{\goodWSig{\Gm}{\tau}{n}} \text{ is defined} \\
  \text{we interpret }\TyLkg{}, \TyTkg{} & \text{ by induction on the input linkage signature, via } \mathit{L} \text{ and } \Glued{\mathit{P}}  \\ 
  \mathit{L}("nil") &=  \denotesC{\top}  \\
  \mathit{L}((A,s,T) :: tl)(\gm)(\gm')(t) &= {\mathit{L}(tl)}(\gm)(\gm')(\lkgproj{1}{t}) \times \Gmodel{\GluedPi{A}{T}}(\gm)(\gm')(\lam{\lkgproj{2}{t}})   \\
  \Glued{\mathit{P}}("nil") &= ( \TyTkg{\LSigEmp} , \denotesC{\top} )\\
  \Glued{\mathit{P}}((A,s,T) :: tl) &= \Glued{\Sigma}(\Glued{\mathit{P}}(tl), \Gsub{T}{\GSubstExt{{\GSubstWeak{1}}}{s}}) \quad \text{(doing substitution on $T$)} \\
  \denotesC{\goodType{\Gm}{\TyLkg{\lsig}}{}} &= {\mathit{L}(\denotesC{\lsig})} \\
  \denotesC{\goodType{\Gm}{\TyTkg{\lsig}}{}} &= \Gmodel{\Glued{\mathit{P}}(\denotesC{\lsig})} \quad \text{ discard syntax info}  \\
  % 
  \denotesC{\goodSig{\Gm}{\LSigEmp}{0}} &= "nil"\\ 
  \denotesC{\goodSig{\Gm}{\LSigAdd{\lsig}{s}{T}}{n+1}} &= (\denotesCC{A}, \denotesCC{s}, \denotesCC{T})::\denotesC{\lsig} \quad \text{ given } \goodTerm{\Gm, \TyTkg{\lsig}}{s}{A} \\ 
  \denotesC{\goodSig{\Gm}{\sub{\lsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
  \denotesC{\goodSig{\Gm}{\lsigproj{1}{\lsig}}{n}} &= "tl"~\denotesC{\lsig}\\ 
  \denotesC{\goodType{\Gm}{\lsigprojT{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[0]} \quad \text{ take the first element in the tuple,...}\\ 
  \denotesC{\goodTerm{\Gm, \TyTkg{\lsigproj{1}{\lsig}}}{\lsigproj{s}{\lsig}}{\lsigprojT{\lsig}}} &= \Gmodel{("hd"~\denotesC{\lsig})[1]} \quad \text{  ... and discard syntax}\\ 
  \denotesC{\goodType{\Gm, \lsigprojT{\lsig}}{\lsigproj{2}{\lsig}}{}} &= \Gmodel{("hd"~\denotesC{\lsig})[2]}\\
  \denotesC{\goodTerm{\Gm}{\LkgEmp}{\TyLkg{\LSigEmp}}} &= \denotesC{\unit} \\
  \denotesC{\goodTerm{\Gm}{\LkgAdd{\lkg}{t}}{\TyLkg{\LSigAdd{\lsig}{s}{T}}}}(\gm)(\gm') &= (\denotesC{\lkg}(\gm)(\gm') , \Gmodel{\Glam{\denotesCC{t}}}(\gm)(\gm')) \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \denotesC{\unit}
  \quad \text{ when \goodSig{\Gm}{\lsig}{0}} \\
  \denotesC{\goodTerm{\Gm}{\Tkg{\lkg}}{\TyTkg{\lsig}}} &= \Gmodel{\Gpair{\Glued{\ell'}}{\Gsub{\Gsub{\denotesCC{\lkgproj{2}{\lkg}}}{\GSubstExt{{\GSubstWeak{1}}}{s}}}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}}}
  \\ &\text{ where } \Glued{\ell'} = \denotesCC{\Tkg{\lkgproj{1}{\lkg}}} 
  \\ &\text{ when } \denotesC{\goodSig{\Gm}{\lsig}{n+1}} = (A, s, T)::\_
  \\ &{~\begin{minipage}{.58\textwidth}\small
    In this definition, $\Gpair{\_}{\_}$ notates the glued interpretation of
    dependent tuple construction, $[\GSubstExt{\_}{\_}]$ notates the glued
    interpretation of substitution extension, and $\Gsub{\_}{\_}$ notates the
    glued interpretation of applying a substitution.
  \end{minipage}}
\end{align*}

%The first complexity comes from the three ``bullets after comma'' notation. The way to interpret this notation is that, this ``$\GSubstExt{\_}{\_}$'' should be considered as a binary function requiring two input of glued interpretation.

%However, a careful reader will notice that, because of the overloading of the notation of dependent tuple construction and substitution extension, our glued notation is also overloaded. The first one ``$\Gpair{\Glued{\ell'}}{...}$'' should be the glued interpretation for dependent tuple, but the other two should be the one for substitution extension.

%Similarly, the second complexity comes from the bullets after the square bracket, which as expected, is the glued interpretation for substitution. Again, ``$\Gsub{\_}{\_}$'' should be considered as a binary function consuming two glued interpretation, corresponding to the substitution operation ``$\sub{\_}{\_}$''.

%Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.

%The right-hand side can be read as follows:
%\begin{enumerate}
%  \item $\denotesCC{\lkgproj{2}{\lkg}}$ is the glued interpretation of a linkage projection
%  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
%  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
%\end{enumerate}
